% We can use verb tag to enter URL but it's better to use native url latex tag
\begin{thebibliography}{99}

\bibitem{sciopps} ``The Long-Baseline Neutrino Experiment: Exploring Fundamental Symmetries of the Universe'',  \url{http://arxiv.org/abs/1307.7335}

\bibitem{cdr_vol2} ``LBNF and DUNE Conceptual Design Report, Vol 2: The Physics Program for DUNE at LBNF'', \url{http://arxiv.org/abs/1512.06148}

\bibitem{cdr_vol4_docdb} LBNE\,DocDB\,10690  % \,\url{http://lbne2-docdb.fnal.gov/cgi-bin/DocumentDatabase/}

\bibitem{surf} Sanford Underground Research Facility: \url{http://www.sanfordlab.org/}

\bibitem{mars} Official MARS site: \url{http://www-ap.fnal.gov/MARS/}

\bibitem{ar39endpoint} LBNE~DocDB~3018% \url{http://lbne2-docdb.fnal.gov:8080/cgi-bin/ShowDocument?docid=3018&version=3}

\bibitem{ar39bkg} ``Measurement of the specific activity of 39Ar in natural argon'', P. Benetti et al. Nucl.Instr. and Meth. A, vol. 574, p. 83, 2007.

\bibitem{atlas1collisions} ``Operating the ATLAS data-flow system with the first LHC collisions'', Journal of Physics: Conference Series 331 (2011) 022007

\bibitem{cms_daq} ``The New CMS DAQ System for Run-2 of the LHC'', IEEE TRANSACTIONS ON NUCLEAR SCIENCE, VOL. 62, NO. 3, JUNE 2015

\bibitem{bnl_protodune_data} \url{https://dune.bnl.gov/wiki/CERN_Prototype_Data_Handling}

\bibitem{slac_rce_1} \url{http://ldrd.slac.stanford.edu/accepted_proposals/2013/Convery.asp}

\bibitem{alice} \url{http://aliceinfo.cern.ch/Public/en/Chapter2/Chap2_TPC.html}

\bibitem{spade_icecube} IceCube\,Data\,Movement, \url{https://icecube.wisc.edu/science/data/datamovement}.

\bibitem{spade_dayabay}"Data processing and storage in the Daya Bay Reactor Antineutrino Experiment" M. He, Nucl. Phys. B Proceedings Supplement 00 (2015) 1–5 (\url{http://arxiv.org/abs/1501.06969}).

\bibitem{fts} A. Norman, M. Diesbug, M. Gheith, R. Illingworth, M. Mengel, ``Archiving Scientific Data Outside of the Traditional HEP Domain, Using the Archive Facilities at Fermilab'', Journal of Physics: Conference Series {\bf 664} (2015) 042039.

\bibitem{atlas_express} "Prompt reconstruction of LHC collision data with the ATLAS reconstruction software" N.Barlow et al, \textit{Journal of Physics: Conference Series 331 (2011) 032004}

\bibitem{pandora} ``The Pandora Software Development Kit for Particle Flow Calorimetry'', Journal of Physics: Conference Series 396 (2012) 022034

\bibitem{muon_bkgd_pdk} ``Muon-induced background to proton decay in the $p \rightarrow K^+\nu$ channel with large underground liquid argon TPC detectors'', \url{http://arxiv.org/abs/1504.06520}

\bibitem{kudr_bkgd_pdk} V.Kudryavtsev, private communication.

\bibitem{sam_chep12} ``Taking Global Scale Data Handling to the Fermilab Intensity Frontier'', Journal of Physics: Conference Series 396 (2012) 032069

\bibitem{dcache_chep12} ``dCache, agile adoption of storage technology'', Journal of Physics: Conference Series 396 (2012) 032077

\bibitem{monarc} \url{http://barone.web.cern.ch/barone/monarc/RCArchitecture.html}

\bibitem{lhc_model_update} \url{http://cds.cern.ch/record/1695401?ln=en}

\bibitem{dirac_wms} \url{http://diracgrid.org/files/docs/index.html}

\bibitem{panda_chep13} ``Evolution of the ATLAS PanDA workload management system for exascale computational science'', Journal of Physics: Conference Series 513 (2014) 032062

\bibitem{glideinwms_chep13} ``Evolution of the pilot infrastructure of CMS: towards a single glideinWMS pool'',  Journal of Physics: Conference Series 513 (2014) 032041

\bibitem{alien} \url{http://alien2.cern.ch/}

\bibitem{jobsub_chep13} ``FIFE -- Jobsub: a grid submission system for intensity frontier experiments at Fermilab'',  Journal of Physics: Conference Series 513 (2014) 032010

\bibitem{osg} \url{http://www.opensciencegrid.org/}

\bibitem{root} ROOT: \url{https://root.cern.ch/}

\bibitem{art} \textit{art}: \url{https://web.fnal.gov/project/ArtDoc/Pages/home.aspx}

\bibitem{geant4} GEANT4: \url{https://geant4.web.cern.ch/geant4/}

\bibitem{jenkins} Jenkins: DUNE\url{https://jenkins-ci.org/}

\bibitem{cvmfs} CVMFS: \url{http://cernvm.cern.ch/portal/filesystem}

\bibitem{duneDocDb1086} DUNE DocDB 1086 %: \mbox{\url{http://docs.dunescience.org:8080/cgi-bin/ShowDocument?docid=1086}}

\end{thebibliography}
