\section{Introduction}
\subsection{DUNE}

The Deep Underground Neutrino Experiment (DUNE) will provide a unique, world-leading program
for the exploration of key questions at the forefront of particle physics and astrophysics~\cite{sciopps,cdr_vol2}.
The two primary detector systems of the DUNE experimental apparatus are:

\begin{enumerate}

\item A high-resolution \textbf{Near Neutrino Detector} just downstream of the neutrino source
\item A massive liquid argon time-projection chamber (LArTPC) deployed as a \textbf{Far
Detector} deep underground and 1,300 km downstream from the source of neutrinos located at FNAL

\end{enumerate}

\noindent
A few options for the DUNE Near Detector are under consideration.  The first of these is a Fine-Grained tracker (FGT) with
an electromagnetic calorimeter and muon identification.  This detector has a channel count of ${\cal O}(10^{5})$.  It is
the reference design, described in the Conceptual Design Report, Vol.~4~\cite{cdr_vol4_docdb}.  The second option is a liquid argon
TPC, to match the Far Detector technology. At the time of writing, a combination of a TPC an a tracker is being
considered as well. The third option is a gaseous-argon TPC.  This document will use the FGT as reference
design, and will be updated as the Near Detector system design evolves.

The Far Detector will be installed at the Sanford Underground Research Facility in Lead, South Dakota~\cite{surf}.
The fiducial mass of the liquid argon will be 40\,kt. In the reference design it is a single-phase LArTPC
with signal readout done with a number of planar arrays of wire electrodes, arranged in stereo pairs.
Approximately 1.5~million readout channels will detect ionization signal resulting from interactions in the detector. 
The Far Detector will also incorporate a photon detection system which will detect scintillation
signal due to ionizing particles traveling within the liquid argon.

The second option for the Far Detector is the dual-phase LArTPC, in which ionization electrons 
drift  in Liquid Argon, reach the surface and are extracted into the gaseous portion of the detector,
due to specially designed electrostatic field configuration. Amplification in the gaseous phase
is achieved, opening possibility for an improved signal to noise ratio.
%This document will focus on the reference (single-phase) design of the Far Detector LArTPC, and will be updated
%as the design evolves.
More details on the physics, design and operation of DUNE detectors 
may be found in the DUNE Conceptual Design Report Vol.4 ~\cite{cdr_vol4_docdb}.

Last but not least, the DUNE Collaboration is engaged in a large-scale  detector prototyping
activity for both LArTPC technology options within the framework of its \textit{\pd} program.
Computing requirements and computing model for \pd is described in a separate series of
documents.
% as detailed in Sec.~\ref{sec:timeline}.


\subsection{About this Document}
\label{sec:modelrole}

This document contains a description of the \textit{Computing Model} (Sec.\ref{sec:computing_model}),
as well as supporting materials that help estimate the parameters of the model and
explain the policies and technical solutions that have been chosen.  These supporting materials are
organized in the following sections:

\begin{itemize}
\item DUNE Data Characteristics (Section~\ref{sec:data-characteristics}).
\item Projected scale and capability of the Far Detector DAQ (Section~\ref{sec:daq}).
\item Near Detector DAQ (Section~\ref{sec:daq-nd})
\item Software and Computing Requirements, Appendix~\ref{sec:requirements}.
\end{itemize}

\noindent
The \textit{Data Characteristics} part describes the characteristics of the data to be collected, processed and otherwise generated in the experiment.
For example, raw data rates generated by DUNE detector systems will determine the scale and required capabilities of the Data
Acquisition System (DAQ).  A brief summary of the Far Detector and Near Detector Data Acquisition is presented in Sections~\ref{sec:daq}
and~\ref{sec:daq-nd} respectively.

The characteristics of the data coming out of DAQ will determine the scale of data storage, network
and other characteristics of the distributed computing system required by DUNE.  Additionally, the scope and projected
volume of Monte Carlo studies  will affect the scale of CPU and storage resources needed by DUNE.  The scope and complexity of the DUNE
reconstruction software impacts both the resources required to process the data and the Monte Carlo, and also affect the
number and size of Monte Carlo studies required, and thus have an impact on the Computing Model. While it is virtually
impossible at the time of writing to provide reliable estimates of the DUNE computing resource requirements a few years out,
the information collected in Section \ref{sec:resource-requirements} provides a basis for extrapolation and order of magnitude
estimations for such resources.

There is a glossary of terms and abbreviations provided in Appendix~\ref{sec:appendix-glossary}. In addition to that,
there are ``Definitions'' sections in the \textit{Requirements} part of the document~(Appendix \ref{sec:requirements}) where
many relevant terms are defined.

The \textit{Requirements}
%which are included as Appendix~\ref{sec:requirements}
define a set of rules and principles by which technologies are selected and specific practices and policies implemented in DUNE computing.
The benefit of having the \textit{Requirements} in place is that they reflect a broad consensus within the Collaboration regarding what it needs from
its computing sector and its components, thus fostering cooperation and better integrations of systems and components in DUNE.

\subsection{Timeline and Evolution of the Computing Model}
\label{sec:timeline}
According to current plans, DUNE will be commissioned in mid-2020s. During the decade before the commissioning
the DUNE Collaboration will continue to be engaged in broad and active R\&D effort, supported by the development of appropriate 
software tools in each area. Extensive simulations are being performed to support design decisions pertaining to the Liquid 
Argon TPC as well as a number of other detector elements and subsystems (cf. the proposed Near Detector, the Photon Detector, 
the target complex etc). Signal processing techniques, background characterization, data reduction, event reconstruction,
and studies of systematic  uncertainties are all under active development. Each of these items impacts the Computing Model
which will naturally evolve as more detailed knowledge and information becomes available.

Importantly, DUNE will continue to be actively involved in the construction and operation of the following LArTPC prototypes:
\begin{itemize}

\item The \textbf{``35t prototype''} --- a Liquid Argon TPC containing 35t of cryogenic medium and situated at FNAL,
that has completed its first cosmic ray run in early 2016.

\item The \textbf{protoDUNE} program, which includes two ``Full Scale'' prototypes of single-phase and dual-phase LArTPCs.
The protoDUNE detectors will be constructed at CERN over the period of time from 2017 to 2018 and tested with a purpose-built
test beam from the CERN SPS. The dual-phase prototype received official CERN experiment designation ``NP02'', and the single-phase ``NP04''.
Both are scheduled for  data taking in 2018.

\end{itemize}

\noindent
Computing plans for the prototypes are presented in Appendix sections~\ref{sec:35t} and \ref{sec:protodune}.
While the prototype program is under way, the DUNE Collaboration will also start preparations for data taking with the
full apparatus, including the Far Detector at the Sanford facility and the Near Neutrino Detector and its associated systems
located at FNAL.  These preparations will involve scaling up the data management and processing systems,
integration testing,  stress testing and realistic data challenges.
The DUNE Computing Model will therefore undergo considerable evolution in the following decade.

This Computing Model is a living document. It first revision was finalized in Fall of 2015.
The rapid evolution of software and computing technologies, combined with the dynamic nature of R\&D in DUNE
and a variety of other factors will require re-evaluation and adjustment of the Computing Model. 
We anticipate the next major update to be produced around 2020.
